Measuring dimension

### 2012-09-04 Dimension/src/dimension.tex
...

\begin{abstract}
\noindent The intrinsic dimension of a dataset is one of the most telling statistics that can be calculated, and one that has received a great deal of attentation over the past decades. Most methods for estimating dimension rely on at least some subjective judgment to make a selection in the data. Adapting methods recently described in \cite{clauset2007power} for the statistical analysis of power laws, we revisit an existing MLE estimator and present a simple, objective procedure to estimate the dimension of a dataset. The procedure runs without human intervention, and produces a value for the dimension, uncertainty bounds and a significance value. We also extend the procedure to extract multiple characteristic dimensions from a dataset and revisit the concept of \textit{lacunarity} in the context of these methods.\end{abstract}
\section{Introduction}
The \textit{intrinsic dimension} of a dataset does not refer to how many features are used to represent it, but---loosely speaking---to how few features could be used to represent it. It is also the exponent of various power laws that determine the way the internal structure of the dataset scales. For many datasets, in particular those with coarse ``fractal'' features or those drawn from a chaotic process, dimension is one of the few well defined statistics available, and usually the most expressive.

@@

For this reason, any scientist who deals with fractal phenomena will want to measure dimension, in order to capture those phenomena in meaningful statistics. Examples are biologists studying the emergence of structure, geologists studying the fractal surfaces and outlines of land masses and rock formations or physicists studying the measures of strange attractors.

@@

Most methods for measuring dimension have relied at least in some part on subjective judgements and processes that make it difficult to capture experimental error. A similar situation was present in the study of power law distributions until recently Clauset, Shalizi and Newman published \cite{clauset2007power}, in which they dealt with various problems, and described a simple objective approach for estimating the exponent of such distributions with no subjective elements and a clear indications of uncertainty and significance with respect of the assumptions.

The methods in this paper come from the realization that the relatively unknown Takens estimator for dimension is also based on the assumption of a power law distribution (albeit of a different form than the ones studied by Clauset et al.), allowing these methods can be adapted to the estimation of dimension. This process also gives us a means to determine multiple dimensions for datasets that can be considered to have different dimensions for different ranges of scales, and we can give a new measure for the lacunarity, a measure of the `gapiness' or texture of a dataset that is often used in combination with dimension.

@@

@@

\subsection{Hausdorff dimension}

[@! is this necessary? Maybe move it to appendix and derive the scaling laws for box counting and correlation there]

@@

The box counting dimension is one of the earlier general-purpose definitions of dimension, and one of the first to be used experimentally. Let $X \subseteq {\mathbb R}^d$. The we can cover the ${\mathbb R}^d$ with a grid of $d$-dimensional hypercubes with sides of length $\epsilon$. We then count how many of these hypercubes contain elements of $X$ and call this value $N_\epsilon$. We assume the following relation:
\begin{align}
\label{box-counting-assumption}
N_\epsilon = \left (\frac{1}{\epsilon}\right) ^ d
\end{align}

@@

In other words a linear relationship in log-log axes. A common approach, especially in the early days of measuring

The logic behind assumption \ref{box-counting-assumption} is detailed in @!. However, this relation is not guaranteed to hold exactly. A common tactic is to assume the relation 
\begin{align}
N_\epsilon = \Phi\left(\frac{1}{\epsilon}\right)\left (\frac{1}{\epsilon}\right) ^ d
\end{align}

Where $\Phi\left(\epsilon\right)$ is called the \textit{pre-factor}. If we say that the pre-factor satisfies
\[
\lim_{\epsilon \rightarrow 0} \frac{\ln \Phi(\frac{1}{\epsilon})}{\ln \frac{1}{\epsilon}} = 0
\]
We can estimate the dimension using 
\[
D_B = \lim_{\epsilon \rightarrow 0} \frac{\ln N_\epsilon}{ \ln \epsilon}
\]

\subsection{Correlation dimension}

Definitions of dimension, fractal dimension and measuring dimension.

Previous work. Theiler. Manifold learning. Databases (Cite faloutsos)We will not discuss the mathematical backgrounds of dimension (Cite Edgar). We will not focus greatly on the algorithmic complexity.

@@

For $q = 2$ we get 
\[
D_2(p) = \lim_{\epsilon \rightarrow 0} \frac{1}{\ln \epsilon}\ln \sum_{c \in C_\epsilon} {p(c)^2}
\]
This is called the correlation dimension, and this is the dimension we focus on. It has the advantage that it is easy to estimate. It should be noted that for many distributions $D_q$ does not vary with $q$, so that we can choose $q$ to suit our needs. Where the dimension does depend on $q$ (the so called \textit{multifractals}) $2$ provides reasonable compromise between caprturing the whole support, and capturing the scaling within the probability density function.

@@

@@

As noted before, we assume that our data is independently sampled from some probability distribution. This may not be the case. In particular, time series data\ldots

[@! Read Theiler's EEG paper. Discuss natural measures briefly. Also ergodicity?]

\section{Methods}

@@

Like the Hill estimator. State assumptions for MLE.

\subsection{Testing the independence assumption}

If we use each point for only one pair, (and our dataset is iid) then the distances for those pairs will be iid samples from $p_D$. If we reuse points, we will invalidate the first assumption. Consider for instance, a dataset of three points $a$, $b$ and $c$ drawn independently from some distribution. If we use only $d(a, b)$ we get a single sample from $p_D$. If we use $d(a, b)$ and $d(a, c)$, the two values are dependent the location of $a$. we are in fact sampling twice independently from the distribution of $d(a, X)$ rather than $d(X, Y)$.

If we use all pairs from a large dataset, however, the recurrence of single points will simply mean that we are sampling from $d(a, Y)$ averaged over many different $a$. Thus we expect that for a large dataset, the distances of all pairs do not form a substantially different distribution than the distances over successive pairs, while giving us a squared increase in the volume of data.

We test this assumption by taking datasets in full (after shuffling the data) and performing the Kolmogorov-Smirnov statistic on them. The results are shown in table @!.

\subsection{Testing the distribution assumption}

\subsection{Statistics and Machine Learning}

\section{Estimating dimension}

\subsection{Maximum Likelihood Estimator}

Derivation of Takens' MLE and its standard error. Comparison to the Hill estimator and discussion of the differences.

@@

Quick review of the methods described in Clauset. Translation of the method to this estimator. Experiment showing the validity of the method for this MLE.

\subsection{Uncertainty}

Quick review of the approach of Clauset.

\subsection{Significance}

Review of the approach of Clauset, and a description of the method translated to this problem.

\subsection{Metric data}

Discussion of the use of dimension for metric data and the use of the significance measure.

@@

@@

\section{Lacunarity}

\section{Experiments}
\subsection{Data sets}

@@

\begin{tabular}{l | r r r r}
\hline
data set & size & dimension & significance & true dimension \\
\hline
Swiss roll & 1000 & & & \\
Noisy swiss roll ! & 1000 & & & \\ 
Sierpinski & 1000 & & & \\
Cantor & & & & \\
Menger Sponge ! & & & & \\
Persistent process ! & & & & \\

@@

British coast line & & & & \\

EEG & & & & \\
JAFFE expression database & & & & \\
Currency (Traina/Faloutsos) ! & & & & \\
Galaxies & & & & \\
DNA random walk & & & & \\
Turning cup & & & & \\ 
Faces & & & & \\ 
Sunspots & & & & \\

@@

@@

\subsection{Derivation of the MLE}

@@

@@

\subsection{Algorithms}

>>>>

\begin{abstract}
\noindent The intrinsic dimension of a dataset is one of the most telling statistics that can be calculated, and one that has received a great deal of attentation over the past decades. Most methods for estimating dimension rely on at least some subjective judgment to make a selection in the data. Adapting methods recently described in \cite{clauset2007power} for the statistical analysis of power laws, we revisit an existing MLE estimator and present a simple, objective procedure to estimate the dimension of a dataset. The procedure runs without human intervention, and produces a value for the dimension, uncertainty bounds and a significance value. We also extend the procedure to extract multiple characteristic dimensions from a dataset and revisit the concept of \textit{lacunarity} in the context of these methods.\end{abstract}
\section{Introduction}
The \textit{intrinsic dimension} of a dataset does not refer to how many features are used to represent it, but---loosely speaking---to how few features could be used to represent it. This is expressed in various power laws that determine how attributes of the datasets scale over varying length scales. For many datasets, in particular those with coarse ``fractal'' features or those generated by deterministic chaos, dimension is one of the few well defined statistics available, and usually the most expressive.

@@

For this reason, any scientist who deals with fractal phenomena will want to measure dimension, in order to capture those phenomena in meaningful statistics. Scientists encoutenering this problem include biologists studying the emergence of structure, geologists studying the fractal surfaces and outlines of land masses and physicists studying the measures of strange attractors.

@@

Most methods for measuring dimension have relied at least in some part on subjective judgements and processes that make it difficult to capture experimental error. A similar situation was present in the study of power law distributions until recently Clauset, Shalizi and Newman published \cite{clauset2007power}, in which they dealt with various problems, and described a simple objective approach for estimating the exponent of such distributions with no subjective elements and a clear indications of uncertainty and significance with respect to the assumptions.

The methods in this paper come from the realization that the relatively unknown Takens estimator for dimension is also based on the assumption of a power law distribution (albeit of a different form than the ones studied by Clauset et al.), allowing these methods to be adapted to the estimation of dimension. This process also gives us a means to determine multiple dimensions for datasets that can be considered to have different dimensions for different ranges of scales, and we can give a new measure for the lacunarity, a measure of the `gapiness' or texture of a dataset that is often used in combination with dimension.

@@

[@! Discuss the range of D for which estimation is reasonable]

@@

\subsection{A theoretical note}

All notions and estimators of dimension described in this paper (and almost all such notions in general) require the data to follow a particular power law, usually between osme length scale and some property of the data. Often this power law can be proved to hold in the limit as the length scale goes to zero, or the power law is show to hold approximately (perhaps for a certain class of sets).

This approach to dimension has two main drawbacks. First, in the limit towards zero for length scales data becomes increasingly sparse . If we draw two points uniformly and measure their distance, it will likely be in the larger part of the range, with probability shrinking rapidly for lower distances. Secondly, we will often also be interested in datasets that seem to have one dimension at some range of scales and another dimension over another range of scales (imagine, for instance, points sampled from a curved line with some small, two-dimensional noise added). In the limit to zero we are likely only measureing the properties of the experimental noise. In the larger distance scales we can measure dimensions that actually reveal information about the processes that shaped our data.

Our solution is not to reduce the meaning of our statistics down to a mathematical definition of dimension, but to take as an assumption that a certain scaling law holds for some range in our data. Data sets, by virtue of filling a geometrical space will generally follow various power laws but they can also deviate from them slightly (see section \ref{lacunarity}). If we simply assume that some scaling law holds exactly, we can take a maximum likelihood estimate of the dimension and use a significance measure to test to what extent our assumption is justified. We can also say that our curved line with added noise, while it technically has dimension 2, behaves as a clean curved line over certain length scales so that in that regime we can expect a 1 dimensional scaling law to hold.

The distinction is a subtle one, and in most cases people will be satisfied to claim simply that they are measuring dimension, but in some situation the distinction may be important.

@@

The box counting dimension is one of the earlier general-purpose definitions of dimension, and one of the first to be used to define an estimator. Let $X \subseteq {\mathbb R}^d$. We can cover the ${\mathbb R}^d$ with a grid of $d$-dimensional hypercubes with sides of length $\epsilon$. We then count how many of these hypercubes contain elements of $X$ and call this value $N_\epsilon$. We assume the following relation:
\begin{align}
\label{box-counting-assumption}
N_\epsilon = \left (\frac{1}{\epsilon}\right) ^ d
\end{align}

@@

In other words a linear relationship in log-log axes. A common approach, especially in the early days of measuring dimension was to use a log-log plot of many values of $N_\epsilon$ and $\epsilon$ and fit a line over a range of $\epsilon$ where the data seems reasonably linear. A more refined approach is to find a range of scales in which the log-log plot looks linear and to perform linear regression on the logarithmes of these values. apart from the subjective element in this approach, linear regression if logarithmic axes has undesirable statistica properties as detailed in \cite{clauset}. [@! Review]

There is no guarantee that this relation holds. A common tactic is to reduce the assumed power law to  
\begin{align}
N_\epsilon = \Phi\left(\frac{1}{\epsilon}\right)\left (\frac{1}{\epsilon}\right) ^ d
\end{align}

where $\Phi\left(\epsilon\right)$ is called the \textit{pre-factor}. If we say that the pre-factor satisfies
\[
\lim_{\epsilon \rightarrow 0} \frac{\ln \Phi(\frac{1}{\epsilon})}{\ln \frac{1}{\epsilon}} = 0
\]
We can estimate the dimension using 
\[
D_B = \lim_{\epsilon \rightarrow 0} \frac{\ln N_\epsilon}{ \ln \epsilon}
\]
This allows many deviations from the power law (such as oscilation) but requires the dimension to be measured in the lower ranges of $\epsilon$ (note that we will need to increse the volume of data for decreasing values of $\epsilon$ if we want to keep getting meaningful measurements).

@@

For $q = 2$ we get 
\[
D_2(p) = \lim_{\epsilon \rightarrow 0} \frac{1}{\ln \epsilon}\ln \sum_{c \in C_\epsilon} {p(c)^2}
\]
This is called the correlation dimension (\ref{correlation dimension}), and this is the dimension we focus on. It has the advantage that it is easy to estimate. It should be noted that for many distributions $D_q$ does not vary with $q$, so that we can choose $q$ to suit our needs. Where the dimension does depend on $q$ (the so called \textit{multifractals}) $2$ provides reasonable compromise between caprturing the whole support, and capturing the scaling within the probability density function.

@@

\subsection{Correlation dimension}
\label{correlation-dimension}
In 1983 Grassberger and Procaccia \cite{grassberger1983measuring} introduced the Correlation dimension. The Correlation dimension is based on the probability that two points, drawn independently wil have a distance less than $r$. We call this probability measure over distances the correlation integral $C(r)$ and we assume the following scaling law:
\begin{align*}
C(r) = \Phi(r) r^D
\end{align*}
Where $Phi(r)$ is again the pre-factor.

The correlation estimator $\widehat{C}(r, X)$ for the correlation integral (over dataset $X$) is straightforward\footnote{$[\ldots]$ are Iverson brackets. They are $1$ when the logical formula within them is true, and $0$ otherwise.}:
\begin{align*}
\widehat{C}(r, X) = \frac{\sum_{x\neq y \in X} [d(x, y) < r]}{\sum_{x \neq y \in X} 1}
\end{align*}

@! Mention the statistical properties of the correlation estimator.

We can substitute $\widehat{C}$ for $C$ and follow the same approach as with the box counting estimator. We plot the $\log C$ against $\log r$, and find the slope of the in an appropriate regime.

The great value of the correlation integral is that it turns a dataset of $n$ points into a dataset of $(n^2 - n)/2$ points. While the set of distances is derived (and thus technically contains no new information), we can treat it as a a set of independent data at very low cost to accuracy (see \ref{independence}). The effect of the correlation integral then, is to give us a polynomial increase in data at almost no cost.

Unfortunately, the Grassberger Procaccia estimtor still suffers form the drawbacks of the box counting estimator. A requirement of a subjective judgement and a reliance on a linear fit in logarithmic axes. The Takens estimator described in section \ref{methods}, does not suffer these problems, so that it what we will base our method on.

@@

As noted before, we assume that our data is independently sampled from some probability distribution. This may not be the case. For instance time series data can be highly autocorrelated. In some cases, for instance when data is generated by a strange attractor, the whole time series may be predetermined by the initial state of the system, so that we cannot speak of a real probability measure, let alone dependency on one.

The solution to the last problem is the natural measure. Let $\{x_i\} \in S$ be an infinite timeseries generated by some process. Then we define the natural measure of that process as
\begin{align*}
p(S') = \lim_{n \rightarrow \infty} \frac{\sum_{x_i, i < n} [xi \in S']}{n}
\end{align*} 
where $S' \subseteq S$ a similar measure can be defined for continuous $i$. In the case of ergodic strange attractors a slight uncertainty about the initial condition of the system will quickly spread out to coincide with the systems natural measure, so that for such systems the natural measure actually functions as a kind of probability measure representing uncertainty of the state of the system after a number of iterations.

Even when we take the natural measure as the measure for which we are measuring the dimension, we have no guarantee that time series data will be independent. In fac, if it were, our time series wouldn't be very interesting. A solution proposed in \cite{theiler1990estimating} is to use for the numerator of the correlation estimator all distances below $r$ except those for which the points are within $w$ time units. The division is still by the number of all point pairs.

A good example of the dangers of autocorrelation in data is provided by \cite{theiler1996re}. For the rest of this paper we will assume independently drawn data.

We should also caution against bootstrapping. In instances drawn form a distribution with infinite support, the probability that the same instance will occur twice in a finite data set is zero so that the distance 0 should never occur. If we boostrap a dataset by sampling with replacement, we will very likely have recurring elements, and thus a number of zero distances. This can lead to a dimension estimate of 0.0 (as would be correct for independent draws with recurring elements, since it suggests a finite support). For this reason if we want to apply the Takens estimator to a bootstrapped dataset, we recommend bootstrapping over the collection of distances rather than the collection of points.

\section{Methods}
\label{methods}

@@

The Takens estimator is based on two assumptions about a dataset $X$ drawn from probability distribution $p$:
\begin{itemize}
  \item The set of distances $R = \{r : x_i, x_j \in X, i>j, d(x_i, x_j)\}$ can be taken as a set of $(n^2 - n)/2$ independent draws from the probability distribution of te random variable $D = d(X, Y)$ where $X$ and $Y$ are independent random variables distributed according to $p$.
  \item The correlation integral follows a power law with a constant pre-factor $C(r) = c r^D_t$ for values of $r$ below some value $r_{\max}$.
\end{itemize}

Under these two assumptions we can derive the following maximum likelihood estimator for $D_t$:
\begin{align*}
\widehat{D_t} &= \frac{|R|}{\sum{R} - |R| \ln r_{\max}}
\end{align*}

The derivation is presented on section \ref{mle-derivation}.

This estimator was first described by Floris Takens in 1985 \cite{takens1985numerical} in relation to measuring dimension, although the estimator for power law distributions (and generic distributions with tail) was already treated in depth by Hill in 1975 \cite{hill1975simple}.

The following assumption discusses the independence asumption and test its influences. The second assumption we take as a hypothesis to be validated by a calculation of the significance (see \ref{significance}).

\subsection{Testing the independence assumption}
\label{independence}
If we use each point for only one pair, (and our dataset is iid) then the distances for those pairs will be iid samples from $p_D$. If we reuse points, we will invalidate the first assumption. Consider for instance, a dataset of three points $a$, $b$ and $c$ drawn independently from some distribution. If we use only $d(a, b)$ we get a single sample from $p_D$. If we use $d(a, b)$ and $d(a, c)$, the two values are dependent on the location of $a$. We are in fact sampling twice independently from the distribution of $d(a, X)$ rather than $d(X, Y)$.

If we use all pairs from a large dataset, however, the recurrence of single points will simply mean that we are sampling from $d(a, Y)$ averaged over many different $a$. Thus we expect that for a large dataset, the distances of all pairs do not form a substantially different distribution than the distances over successive pairs, while giving us a squared increase in the volume of data. In \cite{}

We test this assumption by generating both the set of all distances and the set of distances between successive points (for a shuffled dataset) and calculating the Kolmogorov-Smirnov statistic $k$ for the two datasets. To give us a baseline for comparing the values, we also generate (for the datasets based on generating algorothms) two sets of $n/2$ and $(n^2 - n)/2$ values drawn from the true distribution over distances, and report their KS value $k$. For generated datasets experiments were repeated 100 times and the sample standard deviation is reported in parenthesis after the mean.

The results show that the difference between $k$ and $\widehat{k}$

\begin{tabular}{l | r r r r r r}
\hline
data set & size & features & $\widehat{k}$  & & $k$&  \\
\hline
Swiss roll & 400 & 3 & 0.065 &(0.02) & 0.059&(0.019) \\
Noisy swiss roll & 400 & 3 & 0.068&(0.022)&  0.062&(0.02)\\ 
Sierpinski & 400 & 2 & 0.065&(0.23)& 0.063&(0.19)\\
Cantor & 400 & 1 & 0.06&(0.18) & 0.58&(0.02)\\
Menger Sponge ! & 400 & 3 & & & & \\
Persistent process ! & 400 & 2 & & & & \\

MVN 5 & 400 & 5 & & & & \\

Logistic equation & 400 & 1 & & & & \\
Lorenz attractor & 400 & 3 & & & & \\

British coast line & 400 & 2 & & & & \\   
Expressions & & & & & & \\
Currency ! & & & & & & \\
Galaxies & & & & & & \\
DNA random walk & & & & & & \\
Turning cup & 481 & 245760 & 0.067 & & & \\ 
Faces & & & & & & \\ 
Sunspots & & & & & & \\

Newsgroup posts & & & & & & \\
Protein sequences  & & & & & & \\
Graphs & & & & & & \\
\hline
\end{tabular}

\subsection{Bias and variance of the MLE}

Refer to Smith.

In \cite{smith1992estimating}, Smith derives the following formula for the number of points $N$ (below $d_{\max}$) required to measure the intrinsic dimension to an (mean squared) error of $\nu$, if the intrinsic dimension is $D$:
\begin{align*}
N = D\left[\frac{4D(D+4)}{(D+2)^2}\right]^{\frac{D}{8}} \left[\frac{D+4}{2}\right]^\frac{1}{2} g^\frac{D}{4} \nu^\frac{-(D+4)}{4}
\end{align*}

Where $g$ is a constant depending on the shape of the distribution.

@@

Quick review of the methods described in Clauset. Translation of the method to this estimator.

\subsection{Significance and Uncertainty}
\label{significance}

Following \cite{clauset2007power}, we provide uncertainty values by bootstrapping over the original data: we sample with replacement a dataset the same size as the original and repeat the experiment. We repeat this procedure 100 times and report the sample standard deviation of the outcome.

Since we assume the scaling law to be present in the data and make no claims that for its existence a priori, we must consider the case that we are fitting a distribution to data that does not support it. Again we turn to \cite{clauset2007power} for our approach.

Having fit a power law to the regime $[0, r_{\max}]$ we can generate a dataset that is like our original given that our assumption is true. For the $n$ points in the data w generate from the fitted distribution with probability $\delta$ and sample from the points above $r_{\max}$ with probability $1 - \delta$, where $\delta$ is the proportion of points in the original data below $r_{\max}$. This way we generate a new dataset for the head of the distribution parametrically and the taiol is bootstrapped so that we do not have to make a assumptions about its shape.

We then assume that our data has a bad fit to the distribution and ask what the probability is that such a bad fit would be seen if the assumptions about distribution were true. We generate $n$ datasets by the methods described above, repeat the experiment on them and calculate the KS statistic (note that we calculate the KS statistic of the generated ddata against the distribution fitted on the generated data). The proportion $p$ of trials for which the KS value is above the KS value measured for our original dataset is the probability that such a bad fit would emerge with correct assumptions. Thus, the closer $p$ is to one, the higher our significance (contrary to more traditional significance measures where low values of $p$ indicate).

If $p$ is low (for instance below $0.01$) then the scaling law does not hold for the regime used and we cannot trust the result of the maximum likelihood estimator. In the case of power law distributions the most likely reason for a low $p$ value is that the data fit a distribution that is similar to a power law distribution (which can then be tested by log-likelihood ratio tests). In our case the most likely reason for a low $p$ is hat the power law has a strong non-constant pre-factor which is throwing off the estimator. The cantor set, for instance shows a basic scaling law in log-log axes, but with an oscilating component. In other words the pre-factor is not constant and does not converge to a constant with zero. In these cases the the Takens estimator will fail to converge  as $r_{\max}$ gos to zero (although it oscillates within 0.05 of the true value).

This deviation from the scaling law is called \textit{lacunarity}, and because of the structural properties that sets and measures must posess to deviate from scaling laws, lacunarity is an interesting statistic in itself. We discuss this interpretation further in section \ref{lacunarity}.

@@

For very large datasets, we may run in to the problem that it is not economical to store the set of all distances. We can store the dataset and compute distances on the fly for all elements of the algorithm, except for the KS test, where we require a sorted list of distances. If we are satisfied with a subsample of the set of distances for each run of the KS test, we can then combine this with the method described above to get an approximation to the full estimator with three parameters: the number of candidates, the number of samples per candidate, and the size of the subsample of the data to use for each candidate.

@! Test all three methods

@@

Results are reported in the next section. 
\section{Experiments}
\subsection{Data sets}

@@

\begin{tabular}{l | r r r r}
\hline
data set & size & dimension & significance & true dimension \\
\hline
Swiss roll & 500 & & & \\
Noisy swiss roll ! & 1000 & & & \\ 
Sierpinski & 1000 & & & \\
Cantor & & & & \\
Menger Sponge ! & & & & \\
Persistent process ! & & & & \\

@@

Logistic equation & & & & \\
Lorenz attractor & & & & \\

British coast line & & & & \\   
JAFFE expression database & & & & \\
Currency (Traina/Faloutsos) ! & & & & \\
Galaxies & & & & \\
DNA random walk & & & & \\
Turning cup & & & & \\ 
Faces & & & & \\ 
Sunspots & & & & \\

@@

\section{Lacunarity}
\label{lacunarity}

@@

\subsection{Derivation of the MLE}
\label{mle-derivation}

@@

\subsection{Hausdorff dimension}
\label{hausdorff-dimension}

@@

\subsection{Generating random points}
\subsection{Algorithms}
